Open Preprints in Ecology \& Evolution

### 2012-09-22 ms.tex
starting to integrate all parts into a coherent text

\author{Joel J. Adamson}
\affiliation{Ecology, Evolution and Organismic Biology, University of North Carolina at Chapel Hill}

@@

\keywords{Publishing; arXiv; pre print server, Green Open Access.}

@@

Open preprints servers allow authors to make their manuscripts publicly available before, or in parallel to, submitting them to journals for traditional peer-review. This idea gained popularity 20 years ago with the advent of arXiv, an open preprint server, widely used in the physical sciences \cite{gin11}. The idea behind pre-print servers is fundamentally simple: to make the results of a scientific endavour available to the scientific community as soon as possible rather than wait until the peer-review process is fully completed. The point of arXiv and open preprints servers is not to avoid the peer-review process. Almost all manuscripts on arXiv are submitted to peer-review.  Sharing drafts of research papers on pre-print servers has numerous advantages. These include rapid dissemination of work in progress to a wider audience that includes not only scientists and the general public, but also readers in developing countries where access to subscription journals are often a limiting factor. Further, pre-prints provide the opportunity to solicit feedback from a larger pool of reviewers. As a consequence, pre print servers increase the number of opportunities for review and revision prior to publication, resulting in higher quality submissions which could also alleviate reviewer burden.

In this article, we will first highlight the advantages of open preprints
servers for both scientists and publishers. We will also debunk a few
misconceptions, discuss the policies of major publishers in ecology an
evolution, and briefly review the most popular open preprint servers.

@@

The first and most often discussed advantage of arXiv and open preprints is
speed \ref{fig:map}. The time between submission and the official publication of
a manuscript can be measured in months, sometime in years. For all this time,
the research is only known to a select few: colleagues, editors, reviewers.
Thus, the science cannot be used, discussed, or reviewed by the wider scientific
community. ...

%% Allows papers to be cited earlier (greater immediacy)

% Tim's section:
The review process as a whole is critically over-loaded, because the number of
active scientists increases, because the pressure to publish increases, and
because of an effect dubbed ``the tragedy of the reviewers commons'' REF.  In
the same times, rejection rates are high in most journals (REF), and when the
not invited to submit a revisions, authors are left with the impression that
they must start the whole process all over again. It's thus no surprise that
different initiatives emerged over the last few years, to decrease the time
spent in review. XXX et coll. (REF) called for the recycling and reuse of
peer-reviews: by attaching previous reviews, and detailed replies, to a new
submission, both the editor and the referees can gauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. In a similar way, the
\emph{Peerage of Science} initiative allows authors to seek anonymous pre-
review by their peers. Some journals (LIST?) now accept to publish papers
which received good evaluations, effectively outsourcing the review process. A
widespread use of preprint servers can achieve the same goal of reducing the
time spent in review. By putting a manuscript out there for open comments and
criticisms, the authors will receive valuable feedback, and can improve the
version which will be submitted. With a rich enough community of scientists
depositing preprints, and commenting on them, the process of an open pre-
review can become widespread, and will overall increase the quality of first
submissions.

@@

Preprint servers also establish priority in a fair way.  Some
manuscripts will spend much more time in the review process than
others.  Public preprint servers offer a much fairer way to establish
intellectual priority by making the work available when done, even if
the exact organisation of the manuscript may change. Surprisingly,
there is a perception in biology that public preprints make it easier
to steal ideas, as if scientific ideas only took form in published
material.
% TP This needs more development, I'm not sure I get your point
Mathematicians and physicists have embraced arXiv in part to establish
priority in a fair way\cite{cal12}.

Some of the responses to public preprints are surprising since they are,
essentially, the same as exchanging preprints among colleagues.
% TP Well, no. You trust your colleagues, and you trust that they will
% not give your MS to anyone!
Prepublication reviews by a small network of colleagues is an
important part of the scientific process, which is attested by the
fact that nearly all published papers acknowledge comments by people
not listed as co-authors.  Preprint servers simply offer a way to
extend this network of colleagues to the entire scientific
community. It ensures that science is not constrained by small
networks of scientists exchanging ideas.  Paul Ginsparg created
arXiv.org in part for democratic reasons: he wanted everyone from
students in small universities to Ivy-League professors to have access
to the most recent scientific \emph{ideas}.
% TP One important point to make: the MS and the idea are different things.
%    We should focus more on the later...
Ginsparg's revolutionary idea was simply to use the power of the
internet for preprints, not just for the end product, so the process
can be open from A to Z, instead of being just open at the end of the
process.
% TP Since when is the publishing buisness an open process? :-)
% PhDP: Wishful thinking :) Corrected...

@@

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\ % Victory!
    Science                             	& Accept/Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\ % grrrrrr
    British Ecological Society                  & ? \\ % Waiting for their answer...
    % Oxford worth mentionning?
    \hline
    \end{tabular}
    \caption{Policies for important publishers in ecology and evolution.}
    \label{table:policies}
\end{table*}

There are other ``cultural'' barriers to adoption of preprint servers
by ecologists and evolutionary biologists.  The most notable preprint
server, arXiv, is designed to work best with \TeX{}, a nearly
universal document preparation system among physicists and
mathematicians.  Jackson \cite{jackson2002preprints} further argues
that \TeX{} introduced an ``open source mindset'' to its users, who
now freely share their research findings as well as their software.
Many biologists instead use proprietary software to prepare their
research findings, and many ecology and evolution journals officially
prefer Microsoft Word documents as submissions.  The recent
discipline-wide adoption of free software packages such as R and associated  document generation systems \cite{xie2012}, as well
as interest in open access publishing, has coincided with recent rise in use
of preprint servers by biologists, supporting Jackson's claims.
% Please add this to your citation list: Xie Y (2012). knitr: A general-purpose pack­age for dynamic report gen­er­a­tion in R. R pack­age ver­sion 0.4.1, .

@@

% @Joel
% Rough outline:
% 1. Description of arXiv (origins, funding, maintenance)
% 2. Submission process
% 3. Role within math and physics community (nearly universal), including recent changes
% 4. Nature of citations and "priority"
%
% My biggest problem here is how to cite my sources: the information
% on funding comes from a press release from Cornell University
% Libraries; other basic "raw data" comes straight from the arXiv
% webpages. How do I cite that?

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint
server today, and its use is almost universal in mathematics and most
branches of physics.  Physicist Paul Ginsparg originated arXiv in 1991
for theoretical high-energy physicists to communicate preprints via
email and ftp, and soon thereafter adopted the newly created
world-wide web\cite{jackson2002preprints}.  arXiv now receives over
7,000 submissions per month
(\url{http://arxiv.org/show_monthly_submissions}).  arXiv divides its
submissions into subcategories of physics, mathematics, computer
science, quantitative biology, finance and statistics.  The
quantitative biology category includes subcategories for Populations
and Evolution, Quantitative Methods and other categories that may be
of interest to ecologists and evolutionary biologists.
% Check me on this @Pierre!  Not only is this vague, but I'm not sure
% how to verify it; I want to give people the impression that a
% "Quantitative ecology" or "General ecology" category could be
% created easily
The creation of new categories is regulated by the same system that
categorizes papers.

Submission to arXiv is fully automated via the world-wide web.
Authors can submit \TeX{}/\LaTeX{} documents or directly in PDF or
PostScript (for example, as exported by a word processor).  A
moderation system was put in place in 2004: papers must be categorized
by an ``endorser.''  At least one author of a paper must be an
endorser that has previously submitted a paper or has received
permission to submit to a particular category.  Many authors in
mathematics and physics submit papers as soon as they are ready for
review by colleagues, although another popular option is submitting
simultaneously to a journal and arXiv.

@@

% @PhDP

Figshare (\href{http://figshare.com/}{http://figshare.com/}) is an open servers
that allow scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output: anything

All figshare content (article, figures, datasets) have a unique digital object
identifier (DOI) like any journal article.

All content is licensed under the Creative Commons (CC-BY) license, except
datasets which are published under CC0.
% Small explanation

@@

Posting preprints online increases the community of available informal
peer reviewers, and uses the internet for its original
community-building purposes.  Preprint servers also facilitate
communication between disciplines, bridging cultural as well as
geographic divides.
% JA: perhaps something less advertising-like should go here:
The advantages are clear and the costs are low.

% Karthik's comments: 
% It would be worthwhile to mention that not everyone is stoked about pre-print servers (crazy people). So maybe cite something to that effect?
% ACS Journal Editors’ Policy on Preprints at http://pubs.acs.org/cen/ preprint.% html. Criticisms include: 
% flooding the web with too many repetitive pubs
% premature results without any support from data
% etc.

% BUY NOW!

>>>>

\author{Joel J. Adamson}
\affiliation{Ecology, Evolution and Organismic Biology, University of North
Carolina at Chapel Hill, United-States of America}

@@

\keywords{Publishing; preprint server; arXiv; Green Open Access.}

@@

Public preprints servers allow authors to make their manuscripts publicly
available before, or in parallel to, submitting them to journals for traditional
peer-review. This idea gained popularity 20 years ago with the advent of arXiv,
an open preprint server widely used in the physical sciences \cite{gin11}, but
it also part of other fields' culture. Paul Krugman noted that, in economy, the
\emph{traditional model of submit, get refereed, publish, and then people will
read your work broke down a long time ago. In fact, it had more or less fallen
apart by the early 80s} \cite{kru12}. In addition to a section in arXiv,
economists have also the RePEc (Research Papers in Economics) initiative, which
aims to create an archive of working papers, manuscripts, and book chapters.

The idea behind preprint servers is fundamentally simple: to make the results
available to the scientific community as soon as possible rather than wait until
the peer-review process is fully completed. The point of arXiv and open
preprints servers is not to avoid the peer-review process. Almost all
manuscripts on arXiv are submitted to peer-review.  Sharing drafts of research
papers on preprint servers has numerous advantages. These include rapid
dissemination of work in progress to a wider audience that includes not only
scientists and the general public, but also readers in developing countries
where access to subscription journals are often a limiting factor. Further,
preprints provide the opportunity to solicit feedback from a larger pool of
reviewers. As a consequence, preprint servers increase the number of
opportunities for review and revision prior to publication, resulting in higher
quality submissions which could also alleviate reviewer burden.

Yet, despite the success of the approach in other fields of inquiry, most
manuscripts in ecology and evolution are never submitted to any public archive.
In this article, we will first highlight the advantages of open preprints
servers for both scientists and publishers. We will also debunk a few
misconceptions, discuss the policies of major publishers in ecology an
evolution, and briefly review the most popular open preprint servers.

@@

The first and most often discussed advantage of arXiv and open preprints is
speed \ref{fig:map}. The time between submission and the official publication of
a manuscript can be measured in months, sometime in years. For all this time,
the research is only known to a select few: colleagues, editors, reviewers.
Thus, the science cannot be used, discussed, or reviewed by the wider scientific
community. It is a problem for both scientists and publishers. Manuscripts that
are unknown cannot be integrated and thus take more time to be cited.

The review process as a whole is critically over-loaded, because the number of
active scientists increases, because the pressure to publish increases, and
because of an effect dubbed ``the tragedy of the reviewers commons'' REF.  In
the same times, rejection rates are high in most journals (REF), and when the
not invited to submit a revisions, authors are left with the impression that
they must start the whole process all over again. It's thus no surprise that
different initiatives emerged over the last few years, to decrease the time
spent in review. XXX et coll. (REF) called for the recycling and reuse of
peer-reviews: by attaching previous reviews, and detailed replies, to a new
submission, both the editor and the referees can gauge the work done on the
manuscript, and perhaps evaluate it with less prejudice. In a similar way, the
\emph{Peerage of Science} initiative allows authors to seek anonymous pre-
review by their peers. Some journals (LIST?) now accept to publish papers
which received good evaluations, effectively outsourcing the review process. A
widespread use of preprint servers can achieve the same goal of reducing the
time spent in review. By putting a manuscript out there for open comments and
criticisms, the authors will receive valuable feedback, and can improve the
version which will be submitted. With a rich enough community of scientists
depositing preprints, and commenting on them, the process of an open pre-
review can become widespread, and will overall increase the quality of first
submissions.

@@

Preprint servers also establish priority in a fair way.  Some manuscripts will
spend much more time in the review process than others.  Public preprint servers
offer a much fairer way to establish intellectual priority by making the work
available when done, even if the exact organisation of the manuscript may
change. Surprisingly, there is a perception in biology that public preprints
make it easier to steal ideas, as if scientific ideas only took form in
published material.  Mathematicians and physicists have embraced arXiv in part
to establish priority in a fair way\cite{cal12}.

Some of the responses to public preprints are surprising since they are,
essentially, the same as exchanging preprints among colleagues.  Prepublication
reviews by a small network of colleagues is an important part of the scientific
process, which is attested by the fact that nearly all published papers
acknowledge comments by people not listed as co-authors.  Preprint servers
simply offer a way to extend this network of colleagues to the entire scientific
community. It ensures that science is not constrained by small networks of
scientists exchanging ideas.  Paul Ginsparg created arXiv.org in part for
democratic reasons: he wanted everyone from students in small universities to
Ivy-League professors to have access to the most recent scientific \emph{ideas}.
Ginsparg's revolutionary idea was simply to use the power of the internet for
preprints, not just for the end product, so the process can be open from A to Z,
instead of being just open at the end of the process.
% TP One important point to make: the MS and the idea are different things.  We
% should focus more on the later...

@@

\begin{table*}
    \centering
    \begin{tabular}{|ll|}
    \hline
    Publisher                                   & Policy \\
    \hline
    Springer                            	& Accept \\
    BMC                                 	& Accept \\
    Elsevier                            	& Accept \\
    Nature Publishing Group             	& Accept \\
    Public Library of Science           	& Accept \\
    Royal Society                       	& Accept \\
    National Academy of Science (USA)           & Accept \\
    Ecological Society of America       	& Accept \\
    Oxford Journals                             & Accept \\
    Science                             	& Ambiguous \\
    Wiley-Blackwell                       	& No general policy \\
    British Ecological Society                  & No answer to our query \\
    \hline
    \end{tabular}
    \caption{Policies for important publishers in ecology and evolution.}
    \label{table:policies}
\end{table*}

There are other ``cultural'' barriers to adoption of preprint servers by
ecologists and evolutionary biologists.  The most notable preprint server,
arXiv, is designed to work best with \TeX{}, a nearly universal document
preparation system among physicists and mathematicians.  Jackson
\cite{jackson2002preprints} further argues that \TeX{} introduced an ``open
source mindset'' to its users, who now freely share their research findings as
well as their software.  Many biologists instead use proprietary software to
prepare their research findings, and many ecology and evolution journals
officially prefer Microsoft Word documents as submissions.  The recent
discipline-wide adoption of free software packages such as R and associated
document generation systems \cite{xie2012}, as well as interest in open access
publishing, has coincided with recent rise in use of preprint servers by
biologists, supporting Jackson's claims. \cite{xie12}

@@

% Rough outline:
% 1. Description of arXiv (origins, funding, maintenance)
% 2. Submission process
% 3. Role within math and physics community (nearly universal), including recent changes
% 4. Nature of citations and "priority"
%
% My biggest problem here is how to cite my sources: the information
% on funding comes from a press release from Cornell University
% Libraries; other basic "raw data" comes straight from the arXiv
% webpages. How do I cite that?
% -- Joel

arXiv (\url{http://arxiv.org/}) is the most widely-used preprint server today,
and its use is almost universal in mathematics and most branches of physics.
Physicist Paul Ginsparg originated arXiv in 1991 for theoretical high-energy
physicists to communicate preprints via email and ftp, and soon thereafter
adopted the newly created world-wide web\cite{jackson2002preprints}.  arXiv now
receives over 7,000 submissions per month
(\url{http://arxiv.org/show_monthly_submissions}).  arXiv divides its
submissions into subcategories of physics, mathematics, computer science,
quantitative biology, finance and statistics.  The quantitative biology category
includes subcategories for Populations and Evolution, Quantitative Methods and
other categories that may be of interest to ecologists and evolutionary
biologists.

Submission to arXiv is fully automated via the world-wide web.  Authors can
submit \TeX/\LaTeX documents that are compiled on the server, or directly submit
in PDF/PS format (for example, as exported by a word processor).  A moderation
system was put in place in 2004: papers must be categorized by an ``endorser.''
At least one author of a paper must be an endorser that has previously submitted
a paper or has received permission to submit to a particular category.  Many
authors in mathematics and physics submit papers as soon as they are ready for
review by colleagues, although another popular option is submitting
simultaneously to a journal and arXiv.

@@

Figshare (\href{http://figshare.com/}{http://figshare.com/}) is an open server
that allows scientists to submit any research output: manuscript, figures,
datasets, videos, theses, presentations, and so on. There are no rules to limit
what constitutes a research output and, unlike arXiv, there is no endorser
system (although the administrator reserve the right to remove inappropriate
content). All figshare content (article, figures, datasets) have a unique
digital object identifier (DOI) like any journal article. Tags are used to
classify the content and new tags can be made. All content can be commented and
is licensed under the Creative Commons (CC-BY) license, except datasets which
are published under CC0.

@@

Posting preprints online increases the community of available informal peer
reviewers, and uses the internet for its original community-building purposes.
Preprint servers also facilitate communication between disciplines, bridging
cultural as well as geographic divides. The advantages are clear and the costs are low.

% Short list of disadvantages
